---
title: "Garchinger Heide and restoration sites: <br> Seed mass"
author: "<b>Markus Bauer</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

<br/> <br/> <b>Markus Bauer</b>\*, <b>Sina Appeltauer</b>, <b>Malte Knöppler</b>, <b>Maren Teschauer</b> & <b>Johannes Kollmann</b>

Technichal University of Munich, TUM School of Life Sciences, Chair of
Restoration Ecology, Emil-Ramann-Straße 6, 85354 Freising, Germany

\*[markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

ORCiD ID: [0000-0001-5372-4174](https://orcid.org/0000-0001-5372-4174)
<br>
[Google Scholar](https://scholar.google.de/citations?user=oHhmOkkAAAAJ&hl=de&oi=ao)
<br>
GitHub: [markus1bauer](https://github.com/markus1bauer)
<br>
GitHub: [TUM-Restoration-Ecology](https://github.com/TUM-Restoration-Ecology)

# Preparation

#### Packages

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(ade4)
```

```{r echo = FALSE}
rm(list = ls())
```

#### Load data

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types =
    cols(
      .default = "?",
      treatment = "f",
      treatment_age = "f"
    )
) %>%
  rename(y = CWM_Seed) %>%
  filter(
    !(id %in% c(
      "X2021tum03", "X2021tum27", "X2021tum43", "X2021tum48", "X2021tum51",
      "X2003roederS11", "X2018roederS12", "X2018roederS16", "X2018roederS18"
    )) # Plots with >=10% of Polygonatum odoratum (seed mass = 0.08 g)
  )

sites_fc <- sites %>%
  select(
    id, treatment_age, grass_cover, graminoid_cover
  )

traits <- read_csv(
  here("data", "processed", "data_processed_traits.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types =
    cols(
      .default = "?"
    )
) %>%
  mutate(log_y = log(seedmass)) %>%
  column_to_rownames(var = "accepted_name") %>% 
  select(log_y) %>%
  drop_na() %>%
  rownames_to_column(var = "accepted_name")

species <- read_csv(
  here("data", "processed", "data_processed_species.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types =
    cols(
      .default = "?"
    )
) %>%
  semi_join(traits, by = "accepted_name") %>%
  pivot_longer(-accepted_name, names_to = "id", values_to = "cover") %>%
  pivot_wider(names_from = "accepted_name", values_from = "cover") %>%
  semi_join(sites_fc, by = "id") %>%
  column_to_rownames(var = "id")

traits <- traits %>%
  column_to_rownames(var = "accepted_name")

sites_fc <- sites_fc %>%
  column_to_rownames(var = "id")
```

# Statistics

## Data exploration

### Means and deviations

```{r means}
Rmisc::CI(sites$y, ci = .95)
median(sites$y)
sd(sites$y)
quantile(sites$y, probs = c(0.05, 0.95), na.rm = TRUE)
```

### Graphs of raw data

```{r data-exploration, echo = FALSE, warning = FALSE, message = FALSE}
ggplot(sites, aes(y = y, x = treatment_age)) +
  geom_quasirandom(color = "grey") + geom_boxplot(fill = "transparent") +
  theme(axis.text.x = element_text(angle = 15))
```

### Outliers, zero-inflation, transformations?

```{r outliers, echo = FALSE, warning = FALSE}
sites %>%
  group_by(treatment_age) %>%
  count(treatment)
plot1 <- ggplot(sites, aes(x = treatment_age, y = y)) + geom_quasirandom() +
  theme(axis.text.x = element_text(angle = 15))
plot2 <- ggplot(sites, aes(x = y)) + geom_histogram()
plot3 <- ggplot(sites, aes(x = y)) + geom_density()
plot4 <- ggplot(sites, aes(x = log(y))) + geom_density()
(plot1 + plot2) / (plot3 + plot4)
sites %>%
  filter(y > 5.3)
```

## Model

```{r model-full-dataset, collapse = TRUE}
m <- ade4::fourthcorner(
  tabR = sites_fc, tabL = species, tabQ = traits, modeltype = 6,
  nrepet = 999
)
```

```{r output-model-full-dataset, collapse = TRUE}
summary(m)
m
```

# Subset of dataset

## Set up data

```{r}
subset_sites <- sites %>%
  mutate(
    hay_and_mowing_date = str_c(
      year_hay_transfer, mowing_date, sep = "_"
    )
  ) %>%
  filter(
    is.na(hay_and_mowing_date) | !(hay_and_mowing_date %in% c("1993_summer")),
    is.na(year_topsoil_removal) | year_topsoil_removal != "1996/2003"
  )

subset_sites_fc <- subset_sites %>%
  select(
    id, treatment, grass_cover, graminoid_cover
  ) %>% 
  column_to_rownames(var = "id")

subset_species <- species %>%
  rownames_to_column(var = "id") %>%
  pivot_longer(-id, names_to = "accepted_name", values_to = "cover") %>%
  semi_join(subset_sites, by = "id") %>% 
  pivot_wider(names_from = "accepted_name", values_from = "cover") %>%
  column_to_rownames(var = "id")
```

## Model of subset

```{r model-subset, collapse = TRUE}
m_sub <- ade4::fourthcorner(
  tabR = subset_sites_fc, tabL = subset_species, tabQ = traits, modeltype = 6,
  nrepet = 999
)
```

```{r model-subset-output, collapse = TRUE}
summary(m_sub)
m_sub
```

# Session info

```{r session-info, echo = FALSE}
sessionInfo()
```
